---
title: "MA615 Worldwide Ocean Project"
author: "Team 4: Shuyi Jiang, Jiazhou Liu, Shan Shan, Yihe Tang"
date: "11/6/2017"
output: pdf_document
---

# Region Name: 4

# List of Sources with URLs

###Station 46002 (LLNR 765.1) - WEST OREGON - 275NM West of Coos Bay, OR
###http://www.ndbc.noaa.gov/station_history.php?station=46002
###Station 46028 (LLNR 275) - CAPE SAN MARTIN - 55NM West NW of Morro Bay, CA
###http://www.ndbc.noaa.gov/station_history.php?station=46028
###Station 46089 (LLNR 689) - TILLAMOOK, OR - 85 NM WNW of Tillamook, OR
###http://www.ndbc.noaa.gov/station_history.php?station=46089

# Overview and Summary:
### Report on the condition of the data and any interesting detail
### Discuss the data acquisition, selection of buoys or ships
### Density of data

We carried our data collection of the National Data Buoy Center's Worldwide Ocean Temperature via three locations in North America: #406002, #46028 and #46089, respectively, from 2005 to 2016, within twelve-year span. The baseline and collected factors are: date and time (YYYY-MM-HH-DD), air temperature (ATMP) and water temperature (WTMP) to meet the research goal. In regards to the data per se: the obtained total valid observations are 10839. In addition, the twelve years’ water temperature range is 9.6 &deg;C, with the mean of 13.6 &deg;C; the air temperature range is 15.2&deg;C with the mean value of 12.50&deg;C. For the density of fluctuations, our EDA analysis presents the “noises” regionally and seasonally. The detailed discussions are listed in the body paragraph as follows.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
```

```{r}
library(foreign)
library(ggplot2)
library(car)
library(MASS)
library(arm)
library(dplyr)
library(tidyverse)
library(ggmap)
library(readr)
```

#Point 46028 Data Cleaning 

```{r}
data <- readLines("46028h2005.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2005028 <- read.fwf(textConnection(df), widths=c(14,5,5))
names(data.clean2005028) <- c("YYMMDDHH", "WTMP","ATMP")


```

```{r}
data <- readLines("46028h2006.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2006028 <- read.fwf(textConnection(df), widths=c(14,5,5))
names(data.clean2006028) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46028h2007.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2007028 <- read.fwf(textConnection(df), widths=c(14,5,5))
names(data.clean2007028) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46028h2008.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2008028 <- read.fwf(textConnection(df), widths=c(14,5,5))
names(data.clean2008028) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46028h2009.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 66, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2009028 <- read.fwf(textConnection(df), widths=c(14,6,5))
names(data.clean2009028) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46028h2010.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 66, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2010028 <- read.fwf(textConnection(df), widths=c(14,5,5))
names(data.clean2010028) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46028h2011.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 66, 70), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2011028 <- read.fwf(textConnection(df), widths=c(14,5,5))
names(data.clean2011028) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46028h2012.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2012028 <- read.fwf(textConnection(df), widths=c(14,4,5))
names(data.clean2012028) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46028h2013.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2013028 <- read.fwf(textConnection(df), widths=c(14,4,5))
names(data.clean2013028) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46028h2014.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2014028 <- read.fwf(textConnection(df), widths=c(14,4,5))
names(data.clean2014028) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46028h2015.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 14), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2015028 <- read.fwf(textConnection(df), widths=c(14,5,5))
names(data.clean2015028) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46028h2016.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2016028 <- read.fwf(textConnection(df), widths=c(14,4,5))
names(data.clean2016028) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data.clean46028 <- rbind(data.clean2005028, data.clean2006028, data.clean2007028,
                         data.clean2008028, data.clean2009028, data.clean2010028,
                         data.clean2011028, data.clean2012028, data.clean2013028,
                         data.clean2014028, data.clean2015028, data.clean2016028)
```

```{r}
data46028 <- filter(data.clean46028, substr(data.clean46028$YYMMDDHH,12,13)=="12" )
```

```{r}
data46028$region <- 4
data46028$typereading <- "buoy"
data46028$lat <- "35.712 N"
data46028$lon <- "121.858 W"
data46028$timediff <- "+/- 0"
```

```{r}
data <- data46028[c("region", "typereading", "timediff", "YYMMDDHH",
                    "lat","lon", "WTMP", "ATMP")]
```

```{r}
data2005028 <- filter(data, substr(YYMMDDHH,1,4)==2005)
data2006028 <- filter(data, substr(YYMMDDHH,1,4)==2006)
data2007028 <- filter(data, substr(YYMMDDHH,1,4)==2007)
data2008028 <- filter(data, substr(YYMMDDHH,1,4)==2008)
data2009028 <- filter(data, substr(YYMMDDHH,1,4)==2009)
data2010028 <- filter(data, substr(YYMMDDHH,1,4)==2010)
data2011028 <- filter(data, substr(YYMMDDHH,1,4)==2011)
data2012028 <- filter(data, substr(YYMMDDHH,1,4)==2012)
data2013028 <- filter(data, substr(YYMMDDHH,1,4)==2013)
data2014028 <- filter(data, substr(YYMMDDHH,1,4)==2014)
data2015028 <- filter(data, substr(YYMMDDHH,1,4)==2015)
data2016028 <- filter(data, substr(YYMMDDHH,1,4)==2016)
```

```{r}
data2005028$year <- 2005
data2006028$year <- 2006
data2007028$year <- 2007
data2008028$year <- 2008
data2009028$year <- 2009
data2010028$year <- 2010
data2011028$year <- 2011
data2012028$year <- 2012
data2013028$year <- 2013
data2014028$year <- 2014
data2015028$year <- 2015
data2016028$year <- 2016
```

```{r}
data28 <- rbind(data2005028,data2006028,data2007028,
                data2008028,data2009028,data2010028,
                data2011028,data2012028,data2013028,
                data2014028,data2015028,data2016028)
```

```{r}
data028 <- filter(data28, WTMP != 99.0 & ATMP !=99.0)
data028 <- filter(data028, WTMP != 999.0 & ATMP != 999.0)
data028 <- filter(data028, WTMP != 0 & ATMP != 0)
```
Point 46028 Data Cleaning Finished, chart finished

#Point 46089 Data Cleaning

```{r}
data <- readLines("46089h2005.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2005089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2005089) <- c("YYMMDDHH", "WTMP","ATMP")
```


```{r}
data <- readLines("46089h2006.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2006089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2006089) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46089h2007.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2007089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2007089) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46089h2008.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2008089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2008089) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46089h2009.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2009089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2009089) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46089h2010.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2010089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2010089) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46089h2011.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2011089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2011089) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46089h2012.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2012089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2012089) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46089h2013.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2013089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2013089) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46089h2014.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2014089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2014089) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46089h2015.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2015089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2015089) <- c("YYMMDDHH", "WTMP","ATMP")
```

```{r}
data <- readLines("46089h2016.txt")
df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61, 65))
  df <- rbind(df, subtmp)
}

data.clean2016089 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2016089) <- c("YYMMDDHH", "WTMP","ATMP")
```
# Txt files reading finished
# Join two data frames (datasets) vertically, use the rbind function for data frame combination. 

```{r}
data46089 <- rbind(data.clean2005089,data.clean2006089,data.clean2007089,
                   data.clean2008089,data.clean2009089,data.clean2010089, 
                   data.clean2011089, data.clean2012089, data.clean2013089,
                   data.clean2014089, data.clean2015089,data.clean2016089)
```

# Filter data point at 12 o'clock
```{r}
data46089 <- filter(data46089, substr(data46089$YYMMDDHH,12,13)=="12" )
```

#Add columns
```{r}
data46089$region <- 4
data46089$typereading <- "buoy"
data46089$lat <- "35.712 N"
data46089$lon <- "121.858 W"
data46089$timediff <- "+/- 0"
```

```{r}
data <- data46089[c("region", "typereading", "timediff", "YYMMDDHH",
                    "lat","lon", "WTMP", "ATMP")]
```

#Split By Year
```{r}
data2005089 <- filter(data, substr(YYMMDDHH,1,4)==2005)
data2006089 <- filter(data, substr(YYMMDDHH,1,4)==2006)
data2007089 <- filter(data, substr(YYMMDDHH,1,4)==2007)
data2008089 <- filter(data, substr(YYMMDDHH,1,4)==2008)
data2009089 <- filter(data, substr(YYMMDDHH,1,4)==2009)
data2010089 <- filter(data, substr(YYMMDDHH,1,4)==2010)
data2011089 <- filter(data, substr(YYMMDDHH,1,4)==2011)
data2012089 <- filter(data, substr(YYMMDDHH,1,4)==2012)
data2013089 <- filter(data, substr(YYMMDDHH,1,4)==2013)
data2014089 <- filter(data, substr(YYMMDDHH,1,4)==2014)
data2015089 <- filter(data, substr(YYMMDDHH,1,4)==2015)
data2016089 <- filter(data, substr(YYMMDDHH,1,4)==2016)
```
#Add "year" Column
```{r}
data2005089$year <- 2005
data2006089$year <- 2006
data2007089$year <- 2007
data2008089$year <- 2008
data2009089$year <- 2009
data2010089$year <- 2010
data2011089$year <- 2011
data2012089$year <- 2012
data2013089$year <- 2013
data2014089$year <- 2014
data2015089$year <- 2015
data2016089$year <- 2016
```

```{r}
data46089 <- rbind(data2005089,data2006089,data2007089,
                   data2008089,data2009089,data2010089,
                   data2011089,data2012089,data2013089,
                   data2014089,data2015089,data2016089)
```

```{r}
data089 <- filter(data46089, WTMP != 99.0 & ATMP !=99.0)
data089 <- filter(data089, WTMP != 999.0 & ATMP != 999.0)
data089 <- filter(data089, WTMP != 0 & ATMP != 0)
```
Point 46089 Data Cleaning and Chart Finished

#Point 46002 Data Cleaning
```{r}
#read data files from 2005-2016 at point 46002 
data<-readLines("46002h2005.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61,65))
  df <- rbind(df, subtmp)
}

data.clean2005002 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2005002) <- c("YYMMDDHH","WTMP","ATMP")
```

```{r}
data<-readLines("46002h2006.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61,65))
  df <- rbind(df, subtmp)
}

data.clean2006002 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2006002) <- c("YYMMDDHH","WTMP","ATMP")
```

```{r}
data<-readLines("46002h2007.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61,65))
  df <- rbind(df, subtmp)
}

data.clean2007002 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2007002) <- c("YYMMDDHH","WTMP","ATMP")
```

```{r}
data<-readLines("46002h2008.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61,65))
  df <- rbind(df, subtmp)
}

data.clean2008002 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2008002) <- c("YYMMDDHH","WTMP","ATMP")
```

```{r}
data<-readLines("46002h2009.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61,65))
  df <- rbind(df, subtmp)
}

data.clean2009002 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2009002) <- c("YYMMDDHH","WTMP","ATMP")
```

```{r}
data<-readLines("46002h2011.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61,65))
  df <- rbind(df, subtmp)
}

data.clean2011002 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2011002) <- c("YYMMDDHH","WTMP","ATMP")
```

```{r}
data<-readLines("46002h2012.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61,65))
  df <- rbind(df, subtmp)
}

data.clean2012002 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2012002) <- c("YYMMDDHH","WTMP","ATMP")
```

```{r}
data<-readLines("46002h2013.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61,65))
  df <- rbind(df, subtmp)
}

data.clean2013002 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2013002) <- c("YYMMDDHH","WTMP","ATMP")
```

```{r}
data<-readLines("46002h2014.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61,65))
  df <- rbind(df, subtmp)
}

data.clean2014002 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2014002) <- c("YYMMDDHH","WTMP","ATMP")
```

```{r}
data<-readLines("46002h2015.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61,65))
  df <- rbind(df, subtmp)
}

data.clean2015002 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2015002) <- c("YYMMDDHH","WTMP","ATMP")
```

```{r}
data<-readLines("46002h2016.txt")

df <- NULL
for (i in 2:length(data)) {
  tmp <- data[i]
  subtmp <- paste0(substr(tmp, 1, 13), substr(tmp, 67, 71), substr(tmp, 61,65))
  df <- rbind(df, subtmp)
}

data.clean2016002 <- read.fwf(textConnection(df), widths=c(13,5,5))
names(data.clean2016002) <- c("YYMMDDHH","WTMP","ATMP")
```

```{r}
#bind data files together and add required columns;also we select data only at 12pm
data46002=rbind(data.clean2005002,data.clean2006002,data.clean2007002,
                data.clean2008002,data.clean2009002,data.clean2011002,
                data.clean2012002,data.clean2013002,data.clean2014002,
                data.clean2015002,data.clean2016002,by=c("YYDDMMHH","WTMP","ATMP"))
data46002=data46002%>%
  filter(substr(data46002$YYMMDDHH,12,13)=="12")

data46002$region<-4
data46002$typereading<-"buoy"
data46002$lat<-"42.612N"
data46002$lon<-"130.537W"
data46002$timediff<-"+/- 0"
data<-data46002[c("region","typereading","timediff","YYMMDDHH",
                  "lat","lon","WTMP","ATMP")]
#clean data by removing missing values
data <- filter(data, WTMP != 99.0 & ATMP !=99.0)
data <- filter(data, WTMP != 999.0 & ATMP != 999)
data <- filter(data, WTMP != 0 & ATMP != 0)
```
```{r}
#select data for each year
data2005002 <- filter(data, substr(YYMMDDHH,1,4)==2005)
data2006002 <- filter(data, substr(YYMMDDHH,1,4)==2006)
data2007002 <- filter(data, substr(YYMMDDHH,1,4)==2007)
data2008002 <- filter(data, substr(YYMMDDHH,1,4)==2008)
data2009002 <- filter(data, substr(YYMMDDHH,1,4)==2009)
data2011002 <- filter(data, substr(YYMMDDHH,1,4)==2011)
data2012002 <- filter(data, substr(YYMMDDHH,1,4)==2012)
data2013002 <- filter(data, substr(YYMMDDHH,1,4)==2013)
data2014002 <- filter(data, substr(YYMMDDHH,1,4)==2014)
data2015002 <- filter(data, substr(YYMMDDHH,1,4)==2015)
data2016002 <- filter(data, substr(YYMMDDHH,1,4)==2016)
```

```{r}
#add an extra year column
data2005002$year <- 2005
data2006002$year <- 2006
data2007002$year <- 2007
data2008002$year <- 2008
data2009002$year <- 2009
data2011002$year <- 2011
data2012002$year <- 2012
data2013002$year <- 2013
data2014002$year <- 2014
data2015002$year <- 2015
data2016002$year <- 2016
```

```{r}
#rebind data together
data002 <- rbind(data2005002,data2006002,data2007002,
                 data2008002,data2009002,data2011002,
                 data2012002,data2013002,data2014002,
                 data2015002,data2016002)
```

#46028 EDA
Data Density
```{r}
ggplot(data = data028, aes(x=year, y=WTMP)) + geom_bar(stat = "identity") + 
  labs(title="Volume of Sea Temperature Sample by Year",
       y="Volume of Sample", x="Year")
```

```{r}
ggplot(data = data028, aes(x=year, y=ATMP)) + geom_bar(stat = "identity") + 
  labs(title="Volume of Air Temperature Sample by Year",
       y="Volume of Sample", x="Year")
```

```{r}
#Five Points Summary of Sea Temperature From 2005 To 2016 at Point 46028
summary(data028$WTMP)
```
```{r}
#Table that shows the summary of Sea Temperature for each year
wtmp<- data.frame(matrix(ncol =6, nrow =12))
colnames(wtmp) <- paste0(c("Min","1st Qu","Median","Mean","3rd Qu","Max"), c(1:6))
wtmp[1,]=summary(as.numeric(data2005028$WTMP))
wtmp[2,]=summary(as.numeric(data2006028$WTMP))
wtmp[3,]=summary(as.numeric(data2007028$WTMP))
wtmp[4,]=summary(as.numeric(data2008028$WTMP))
wtmp[5,]=summary(as.numeric(data2009028$WTMP))
wtmp[6,]=summary(as.numeric(data2010028$WTMP))
wtmp[7,]=summary(as.numeric(data2011028$WTMP))
wtmp[8,]=summary(as.numeric(data2012028$WTMP))
wtmp[9,]=summary(as.numeric(data2013028$WTMP))
wtmp[10,]=summary(as.numeric(data2014028$WTMP))
wtmp[11,]=summary(as.numeric(data2015028$WTMP))
wtmp[12,]=summary(as.numeric(data2016028$WTMP))
wtmp$year<-c(2005,2006,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016)
wtmp<-wtmp[c(7,1,2,3,4,5,6)]
wtmp
```


```{r}
#Boxplot of Sea Temperature from 2005 to 2016 at Point 46028
ggplot(data028,aes(x=year, y=WTMP,group=year, fill=year))+geom_boxplot() +
  labs(title="Sea Temperature from 2005 to 2016 at Point 46028",
       y="Sea Temperature", x="Year")
```
Year 2015 has seen the highest sea temperature compare to other years.
```{r}
#Five Points Summary of Air Temperature From 2005 To 2016 at Point 46028
summary(data028$ATMP)
```
```{r}
#Table that shows the summary of Air Temperature for each year
atmp<- data.frame(matrix(ncol =6, nrow =12))
colnames(atmp) <- paste0(c("Min","1st Qu","Median","Mean","3rd Qu","Max"), c(1:6))
atmp[1,]=summary(as.numeric(data2005028$ATMP))
atmp[2,]=summary(as.numeric(data2006028$ATMP))
atmp[3,]=summary(as.numeric(data2007028$ATMP))
atmp[4,]=summary(as.numeric(data2008028$ATMP))
atmp[5,]=summary(as.numeric(data2009028$ATMP))
atmp[6,]=summary(as.numeric(data2010028$ATMP))
atmp[7,]=summary(as.numeric(data2011028$ATMP))
atmp[8,]=summary(as.numeric(data2012028$ATMP))
atmp[9,]=summary(as.numeric(data2013028$ATMP))
atmp[10,]=summary(as.numeric(data2014028$ATMP))
atmp[11,]=summary(as.numeric(data2015028$ATMP))
atmp[12,]=summary(as.numeric(data2016028$ATMP))
atmp$year<-c(2005,2006,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016)
atmp<-atmp[c(7,1,2,3,4,5,6)]
atmp
```

```{r}
#Boxplot of Air Temperature from 2005 to 2016 at Point 46028
ggplot(data028,aes(x=year, y=ATMP,group=year, fill=year))+geom_boxplot() +
  labs(title="Air Temperature from 2005 to 2016 at Point 46028",
       y="Aea Temperature", x="Year")
```
Year 2015 has seen the highest air temperature, and the tendency is similar to the sea temperature graph. 
```{r}
#Lineplot of Air temperatue from 2005 to 2016 at Point 46028
ggplot(data028,aes(x=as.numeric(row.names(data028)),y=as.numeric(ATMP)))+
  geom_line()+labs(title="Air Temperature from 2005 to 2016 at Point 46028",
                   y="Air Temperature", x="Days")
```
Air temperature from 2005 to 2016 at point 46028 slightly rised. 
```{r}
#Lineplot of Sea Temperatue from 2005 to 2016 at Point 46028
ggplot(data028,aes(x=as.numeric(row.names(data028)),y=as.numeric(WTMP)))+
  geom_line()+labs(title="Sea Temperature from 2005 to 2016 at Point 46028", 
                   y="Sea Temperature", x="Days")
```
Sea temperature from 2005 to 2016 slightly rised in the last couple of years.
#EDA of Point 46089
Data Density
```{r}
ggplot(data = data089, aes(x=year, y=WTMP)) + geom_bar(stat = "identity") + 
  labs(title="Volume of Sea Temperature Sample by Year",
       y="Volume of Sample", x="Year")
```

```{r}
ggplot(data = data089, aes(x=year, y=ATMP)) + geom_bar(stat = "identity") + 
  labs(title="Volume of Air Temperature Sample by Year", 
       y="Volume of Sample", x="Year")
```
Year 2006, 2007, and 2008 have much less data than other years.
```{r}
#Five Points Summary of Sea Temperature From 2005 To 2016 at Point 46089
summary(data089$WTMP)
```

```{r}
#Table that shows the summary of Sea Temperature for each year
wtmp<- data.frame(matrix(ncol =6, nrow =12))
colnames(wtmp) <- paste0(c("Min","1st Qu","Median","Mean","3rd Qu","Max"), c(1:6))
wtmp[1,]=summary(as.numeric(data2005089$WTMP))
wtmp[2,]=summary(as.numeric(data2006089$WTMP))
wtmp[3,]=summary(as.numeric(data2007089$WTMP))
wtmp[4,]=summary(as.numeric(data2008089$WTMP))
wtmp[5,]=summary(as.numeric(data2009089$WTMP))
wtmp[6,]=summary(as.numeric(data2010089$WTMP))
wtmp[7,]=summary(as.numeric(data2011089$WTMP))
wtmp[8,]=summary(as.numeric(data2012089$WTMP))
wtmp[9,]=summary(as.numeric(data2013089$WTMP))
wtmp[10,]=summary(as.numeric(data2014089$WTMP))
wtmp[11,]=summary(as.numeric(data2015089$WTMP))
wtmp[12,]=summary(as.numeric(data2016089$WTMP))
wtmp$year<-c(2005,2006,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016)
wtmp<-wtmp[c(7,1,2,3,4,5,6)]
wtmp
```


```{r}
#Boxplot of Sea Temperature from 2005 to 2016 at Point 46089
ggplot(data089,aes(x=year, y=WTMP,group=year, fill=year))+geom_boxplot() +
  labs(title="Sea Temperature from 2005 to 2016 at Point 46089",
       y="Sea Temperature", x="Year")
```
Sea temperature at 46089 at year 2006 is less various than other years. 
```{r}
#Five Points Summary of Air Temperature From 2005 To 2016 at Point 46089
summary(data089$ATMP)
```
```{r}
#Table that shows the summary of Air Temperature for each year
atmp<- data.frame(matrix(ncol =6, nrow =12))
colnames(atmp) <- paste0(c("Min","1st Qu","Median","Mean","3rd Qu","Max"), c(1:6))
atmp[1,]=summary(as.numeric(data2005089$ATMP))
atmp[2,]=summary(as.numeric(data2006089$ATMP))
atmp[3,]=summary(as.numeric(data2007089$ATMP))
atmp[4,]=summary(as.numeric(data2008089$ATMP))
atmp[5,]=summary(as.numeric(data2009089$ATMP))
atmp[6,]=summary(as.numeric(data2010089$ATMP))
atmp[7,]=summary(as.numeric(data2011089$ATMP))
atmp[8,]=summary(as.numeric(data2012089$ATMP))
atmp[9,]=summary(as.numeric(data2013089$ATMP))
atmp[10,]=summary(as.numeric(data2014089$ATMP))
atmp[11,]=summary(as.numeric(data2015089$ATMP))
atmp[12,]=summary(as.numeric(data2016089$ATMP))
atmp$year<-c(2005,2006,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016)
atmp<-atmp[c(7,1,2,3,4,5,6)]
atmp
```

```{r}
#Boxplot of Air Temperature from 2005 to 2016 at Point 46089
ggplot(data089,aes(x=year, y=ATMP,group=year, fill=year))+geom_boxplot() + 
  labs(title="Air Temperature from 2005 to 2016 at Point 46089", 
       y="Aea Temperature", x="Year")
```
Year 2008 has seen the highest average temperature from 2005 to 2016
```{r}
#Lineplot of Air temperatue from 2005 to 2016 at Point 46089
ggplot(data089,aes(x=as.numeric(row.names(data089)),y=as.numeric(ATMP)))+
  geom_line()+labs(title="Air Temperature from 2007 to 2016 at Point 46089",
                   y="Air Temperature", x="Days")
```
Air temperature fluctuated during the ten year period, whereas sea temperature in the next graph is more stable. 


```{r}
#Lineplot of Sea Temperatue from 2005 to 2016 at Point 46089
ggplot(data089,aes(x=as.numeric(row.names(data089)),y=as.numeric(WTMP)))+
  geom_line()+labs(title="Sea Temperature from 2007 to 2016 at Point 46089",
                   y="Sea Temperature", x="Days")
```
Sea Temperature at point 46089 was stable from 2005 to 2016. 
#EDA of Point 46002
Data Density
```{r}
ggplot(data = data002, aes(x=year, y=as.numeric(WTMP))) + geom_bar(stat = "identity") +
  labs(title="Volume of Sea Temperature Sample by Year", 
       y="Volume of Sample", x="Year")
```

```{r}
ggplot(data = data002, aes(x=year, y=as.numeric(ATMP))) + geom_bar(stat = "identity") +
  labs(title="Volume of Air Temperature Sample by Year", 
       y="Volume of Sample", x="Year")
```
Data in 2010 is missing. 
```{r}
#Table that shows the summary of Air Temperature for each year
atmp<- data.frame(matrix(ncol =6, nrow =11))
colnames(atmp) <- paste0(c("Min","1st Qu","Median","Mean","3rd Qu","Max"), c(1:6))
atmp[1,]=summary(as.numeric(data2005002$ATMP))
atmp[2,]=summary(as.numeric(data2006002$ATMP))
atmp[3,]=summary(as.numeric(data2007002$ATMP))
atmp[4,]=summary(as.numeric(data2008002$ATMP))
atmp[5,]=summary(as.numeric(data2009002$ATMP))
atmp[6,]=summary(as.numeric(data2011002$ATMP))
atmp[7,]=summary(as.numeric(data2012002$ATMP))
atmp[8,]=summary(as.numeric(data2013002$ATMP))
atmp[9,]=summary(as.numeric(data2014002$ATMP))
atmp[10,]=summary(as.numeric(data2015002$ATMP))
atmp[11,]=summary(as.numeric(data2016002$ATMP))
atmp$year<-c(2005,2006,2007,2008,2009,2011,2012,2013,2014,2015,2016)
atmp<-atmp[c(7,1,2,3,4,5,6)]
atmp
```

```{r}
#Table that shows the summary of Sea Temperature for each year
wtmp<- data.frame(matrix(ncol =6, nrow =11))
colnames(wtmp) <- paste0(c("Min","1st Qu","Median","Mean","3rd Qu","Max"), c(1:6))
wtmp[1,]=summary(as.numeric(data2005002$WTMP))
wtmp[2,]=summary(as.numeric(data2006002$WTMP))
wtmp[3,]=summary(as.numeric(data2007002$WTMP))
wtmp[4,]=summary(as.numeric(data2008002$WTMP))
wtmp[5,]=summary(as.numeric(data2009002$WTMP))
wtmp[6,]=summary(as.numeric(data2011002$WTMP))
wtmp[7,]=summary(as.numeric(data2012002$WTMP))
wtmp[8,]=summary(as.numeric(data2013002$WTMP))
wtmp[9,]=summary(as.numeric(data2014002$WTMP))
wtmp[10,]=summary(as.numeric(data2015002$WTMP))
wtmp[11,]=summary(as.numeric(data2016002$WTMP))
wtmp$year<-c(2005,2006,2007,2008,2009,2011,2012,2013,2014,2015,2016)
wtmp<-wtmp[c(7,1,2,3,4,5,6)]
wtmp
```

```{r}
#Boxplot of Air Temperatue from 2005 to 2016 at Point 46002
ggplot(data002,aes(x=year, y=as.numeric(ATMP),group=year, fill=year))+geom_boxplot() +
  labs(title="Air Temperature from 2007 to 2016 at Point 46002",
       y="Air Temperature", x="Year")
```

```{r}
#Boxplot of Sea Temperatue from 2005 to 2016 at Point 46002
ggplot(data002,aes(x=year, y=as.numeric(WTMP),group=year, fill=year))+geom_boxplot() +
  labs(title="Sea Temperature from 2007 to 2016 of Point 46002",
       y="Sea Temperature", x="Year")
```
The tendency of air and sea temperature at point 46002 are similar throughout 10 years period. 
```{r}
#Lineplot of Air temperatue from 2005 to 2016 at Point 46002
ggplot(data002,aes(x=as.numeric(row.names(data002)),y=as.numeric(ATMP)))+
  geom_line()+labs(title="Air Temperature from 2007 to 2016 at Point 46002",
                   y="Air Temperature", x="Days")
```
Air temperature at point 46002 fluctuate every year and slightly rises from year 2005 to 2016. 

```{r}
#Lineplot of Sea Temperatue from 2005 to 2016 at Point 46002
ggplot(data002,aes(x=as.numeric(row.names(data002)),y=as.numeric(WTMP)))+
  geom_line()+labs(title="Sea Temperature from 2007 to 2016 at Point 46002",
                   y="Sea Temperature", x="Days")
```
Sea temperature at point 46002 fluctuate every year and slightly rises from year 2005 to 2016. 

#Three Buoy Points Summary
```{r}
#Bind Three Points Together

data028$point <- "46028"
data002$point <- "46002"
data089$point <- "46089"
data_all <- rbind(data002,data028,data089)
```

```{r}
#Comparision of Air Temperature of the Three Points from 2005 to 2016
ggplot(data_all,aes(x=as.numeric(row.names(data_all)),y=as.numeric(ATMP),
                    group=point,color=point))+geom_line()+
  labs(title="Air Temperature of the Three Points from 2005 to 2016", 
       y="Air Temperature", x="Days")
```
Point 46002 and 46089, which are closer to the continent, where the air temperature varies more throughout years than point 46028, which is farther away from the continent.

```{r}
#Comparision of Sea Temperature of the Three Points from 2005 to 2016
ggplot(data_all,aes(x=as.numeric(row.names(data_all)),y=as.numeric(WTMP),
                    group=point,color=point))+geom_line()+
  labs(title="Sea Temperature of the Three Points from 2005 to 2016",
       y="Air Temperature", x="Days")
```
Similarly, point 46002 and 46089, which are closer to the continent, where the sea temperature varies more throughout years than point 46028, which is farther away from the continent.

```{r}
#Map
data_map<-read.csv("Data_map.csv",header=TRUE)
map <- ggmap(get_map(location =c(-130,45),scale =1, zoom = 3, 
                     source = "google", maptype = "roadmap")) +
  geom_point(aes(x=Longitude,y=Latitude),                                                                    data=data_map,alpha=.5,size = 1,color="red")
  scale_size(range=c(3,20))

map
```
